---
title: "ACTIVITY 4"
output:
  html_notebook: default
  pdf_document:
    latex_engine: xelatex
  html_document: default
---

## Section A

Load "iqphys.csv" and do a multiple linear regression:

- Which of the predictors explains some of the variability in intelligence quotient (IQ)?
- What is the effect of brain size (Brain), after taking into account weight and height?
- Give a confidence interval for the coefficients
- What is the IQ confidence interval for the predictor values: brain size = 95, height = 70, weight = 180?
- Do the residuals follow a normal distribution? Make the histogram and quantile-quantile plot representations convenient.

## Section B

Load the `prostate.csv` dataset
- Remove the train column
- What is the best model that explains the variable lpsa?
- Give a confidence interval for the coefficients


## Section C

Loads the Wisconsin ICU breast cancer dataset ("wisconsin_breast.csv").

Divide it into train - 400 first rows and test - row 401 onwards.

```{r}
df <- read.csv("wisconsin_breast.csv")
dim(df)
head(df)
str(df)
```


We set the target variable (V2) as well as the train and test sets

```{r}
df$V2 <- as.numeric(as.factor(df$V2) ) - 1
df_train <- df[1:400, 1:32]
df_test <- df[401:569, 1:32]

X_train <- df_train[,4:32]
y_train <- df_train[,3]

X_test <- df_test[,4:32]
y_test <- df_test[,3]
```

Do a logistic regression making model selection according to AIC and giving the confusion table and metrics in the test:

- What variables are most influential in having breast cancer?
- Select the model using StepAIC and make the prediction on the test set giving the confusion matrix.

Better results are understood in this case as the best **Recall** by not wanting to send sick patients home.
